---
title: "Logistic Regression"
author: "Samet Ozturk"
date: "11/21/2019"
output:
  html_document: default
  pdf_document: default
  word_document: default
---
# DATA
UIS Drug Treatment study data
Description
There are 628 data points in the original data, 575 of which have no missing values.

Variable descriptions:

Variable_____Description___________________________Codes/Values

ID__________Identification Code________________________ 1 - 628

AGE_________Age at Enrollment	Years

BECK________Beck DepressionScore____________________0.000 - 54.000

IV__________History of IV Drug Use________________1 = Never_________2 = Previous_________3 = Recent

NDRUGTXT_____Number of Prior Drug Treatments______________0 - 40

RACE________Subject's Race_________0 = White_________1 = Non-White

TREAT_______Treatment Randomization Assignment______0 = Short____1 = Long

SITE________Treatment Site_______________________0 = A_________1 = B

DFREE_______Remained Drug Free for 12 Months________1 = Yes_________0 = No

Usage
data(uis)
Format
A data frame with dimension 575 by 18.

Source
Table 1.3 of Hosmer,D.W. and Lemeshow, S. (1998)

References
Hosmer,D.W. and Lemeshow, S. (1998) Applied Survival Analysis: Regression Modeling of Time to Event Data, John Wiley and Sons Inc., New York, NY

## MODELLING
We will load data and check the structure.
```{r}
df_logreg <- read.table(file = "uis.dat", header = T)
str(df_logreg)
```
#### We will take DFREE as our response variable and define a binomial logistic regression model. We will defince categorical variables with more than 2 levels as factor. Then validate our model, testing our hypothesis with %5 significance level.
```{r}
logreg1 <- glm(DFREE~AGE+BECK+as.factor(IVHX)+NDRUGTX+RACE+TREAT+SITE,family = binomial, data = df_logreg)
summary(logreg1)
ResidualDeviance = 619.25
dev_dif <- logreg1$null.deviance - ResidualDeviance
dev_dif
```
Difference between null deviance and residual deviance should be greater than chi-square table value. We will find chi-square table value with 8 degrees of freedom(because of our feature number in our model.) and with 0.05.
table value: 15.507
Our model is significant.
For Chi-Squared Table: <https://encrypted-tbn0.gstatic.com/images?q=tbn%3AANd9GcRO7PIeI6aFiPDs6Qwj_GH0rKuqkqJzIiCHTUO9EpwZAMqFKWY0>




#### Interpretation of AGE and RACE coefficients:
We will use exp() function because model summary gives us the coefficients for log(odds) of DFREE.(Remember logit and sigmoid functions.)

```{r}
exp(logreg1$coefficients)
```
When AGE increseas for one unit(or compared to one year younger, on year older person), remaining drug free for 12 months odds increases 1.05 times.

In RACE feature, 1 represent for being white, 0 represent not being black.
Compared to other races white person's remaining drug free for 12 months odds is 1.25360854 times.

#### Now let's do our hypothesis test for significance(%5 significance level) of features by interpreting p-values.

AGE
H0: ß1 = 0
H1: ß1 != 0
p = 5.59e-05     Rejecting null hypothesis. Feature is statistically significant.

BECK
H0: ß2 = 0
H1: ß2 != 0
p = 0.97961     Can not reject null hypothesis. Feature is not statistically significant.

as.factor(IVHX)2 
H0: ß3 = 0
H1: ß3 != 0
p = 0.03581      Rejecting null hypothesis. Feature is statistically significant.

as.factor(IVHX)3
H0: ß4 = 0
H1: ß4 != 0
p = 0.00401     Rejecting null hypothesis. Feature is statistically significant.

NDRUGTX
H0: ß5 = 0
H1: ß5 != 0
p = 0.01642     Rejecting null hypothesis. Feature is statistically significant.

RACE
H0: ß6 = 0
H1: ß6 != 0
p = 0.31159    Can not reject null hypothesis. Feature is not statistically significant.

TREAT
H0: ß7 = 0
H1: ß7 != 0
p = 0.02640     Rejecting null hypothesis. Feature is statistically significant.

SITE
H0: ß8 = 0
H1: ß8 != 0
p = 0.49375     Can not reject null hypothesis. Feature is not statistically significant.

#### We will compute Confidence Intervals(with %95 confidence) for odds ratio.

```{r}
exp(confint(logreg1,level = 0.95))
```
When AGE increseas for one unit(or compared to one year younger, on year older person), remaining drug free for 12 months odds increases between 1.0165999 and 1.0885461 with %95 confidence level.

When BECK depression score increseas for one unit(or compared to one unit lower person), remaining drug free for 12 months odds increases between 0.9791804 and 1.0216018 with %95 confidence level.

Previous IV drug users' remaining drug free for 12 months odds higher between
0.3071122 and 0.9511436 compared to others with %95 confidence level.

Recent IV drug users' remaining drug free for 12 months odds higher between
0.2894769 and 0.7879531 compared to others with %95 confidence level.

When number of prior drug treatments increases by one unit remaining drug free for 12 months odds increase is between 0.8909971 and 0.9854565 times with %95 confidence level.

Compared to other races white person's remaining drug free for 12 months odds is between 0.8050679 and 1.9351600.
-------
Long term treatment group's remaining drug free for 12 months odds is higher between 1.0550552 and 2.3068471 times compared to short term treatment group with %95 confidence level.

B settlement treatment group's remaining drug free for 12 months odds is higher between 1.0550552 and 2.3068471 times compared to A settlement treatment group with %95 confidence level.


#### We will select our final model with "Stepwise Regression" algorithm.
```{r}
step(glm(DFREE~AGE+BECK+factor(IVHX)+NDRUGTX+RACE+TREAT+SITE,data=df_logreg),direction="both")
```

Our final model is:
```{r}
logreg_final <- glm(DFREE~AGE+factor(IVHX)+NDRUGTX+TREAT, data = df_logreg, family = binomial)
summary(logreg_final)
```

#### We will find specificity and sensitivity values for our final model. We will use response variable's mean for cut point(or threshold) instead of default 0.5 value.
```{r}
cut_point <- mean(df_logreg$DFREE)
p.pred.dfree<-fitted(logreg_final)  
p.pred.dfree[p.pred.dfree>cut_point]=1
p.pred.dfree[p.pred.dfree<=cut_point]=0
pred_dfree <- as.factor(p.pred.dfree)
DFREE <- as.factor(df_logreg$DFREE)

sens.dfree<-caret::sensitivity(pred_dfree,DFREE,positive=levels(DFREE)[2])
spec.dfree<-caret::specificity(pred_dfree,DFREE,negative=levels(DFREE)[1])

paste("sensitivity :",sens.dfree, sep = " ")
paste("specificity :", spec.dfree, sep = " ")

```
* TP: True Positive
* FP: False Positive
* TN: True Negative
* FN: False Negative

Sensivitiy = TP/(TP+FN) = %65.3
On all true remaining drug-free persons(for 12-month period), our model predicted %65.3 correctly. 
In other words, our model tagged %34.7 truly remained drug-free persons as "not remained drug-free for 12 months".


Specificity = TN/(FP+TN) = % 60.9
On all truly not remained drug free persons(for 12-month period) our model could label those persons correctly with %60.9 accuracy.
In other words, our model tagged %39.1 of people as remained drug-free while they did not.


#### Now let's create a new observation(or define a new person) to do prediction. 
Note: Normally we do this by splitting our data set via splitting into two as train and test. But for the sake of simplicity and understandability we won't do train-test split and cross validation.

We are creating a person with age of 30, beck depression score 9.01, history of IV drug use is 2, number of previous drug treatment is 1, race is white, short treatment group, settlement A and real remaning drug free value is no.
```{r}
newdata <- data.frame("AGE"= 30, "BECK"= 9.01, "IVHX"= 2, "NDRUGTX"= 1, "RACE"=1, "TREAT" = 0, "SITE" = 0, "DFREE" = 0 )
predict(logreg_final,newdata = newdata)

```
Remember again, this is the log(odds) ratio. To find real odd ratio, we will use exp(which means e^y) to get rid of log.
```{r}
exp(predict(logreg_final,newdata = newdata))
```
This person's remaining drug free odds is 0.23 which is consistent with our real value.


